---
title: "Get started with pkgmkr"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgmkr)
```

# Start here

`pkgmkr` helps you create a new R package without doing the same setup by hand
every time. 

It can be used for a quick one-time setup or with a config file you
reuse for future packages.

## Create a package with one function call

Start with `mk_pkg()` if you want a new package with the main setup files.

```{r eval = FALSE}
library(pkgmkr)

mk_pkg(
  path = "mypackage",
  author = "Jane Doe",
  email = "jane.doe@example.com",
  git = FALSE,
  check_pkg_name = FALSE,
  pkgdown = FALSE
)
```

This creates a package called `mypackage` in your current working directory.

### What the main arguments do

- `path` sets the package folder name or full path.
- `author` is the author name. It must include a first and last name.
- `email` is optional. If you include it, it is added to the package metadata.
- `git` controls whether a Git repository is initialized.
- `check_pkg_name` controls whether `pkgmkr` checks if the package name is
  available using the `available` package.
- `pkgdown` controls whether a pkgdown site is set up and built.

## What gets created when you run `mk_pkg()`

`pkgmkr` handles the basic setup so you can start writing code sooner.

After a successful run, your package will include the main files you would
usually make by hand:

- `DESCRIPTION` for package metadata
- `NAMESPACE` for exported functions
- `R/` for your package code
- `man/` for generated documentation
- license files based on your selected license
- `README.md` if `readme_md = TRUE`
- Git setup if `git = TRUE`
- pkgdown setup if `pkgdown = TRUE`

## Use options when you need more control

You do not need every argument the first time. Start small, then add options
when you need them.

For example, if you want Git and pkgdown from the start:

```{r eval = FALSE}
mk_pkg(
  path = "~/projects/myanalysispkg",
  author = "Jane Doe",
  email = "jane.doe@example.com",
  git = TRUE,
  git_username = "janedoe",
  git_email = "jane.doe@example.com",
  license = "GPL-3",
  pkgdown = TRUE
)
```

Use this when you already want Git and a pkgdown site.

## Create a reusable package template with a config file

If you make packages often, keep one config file with your usual settings. Then
you can reuse it instead of typing the same values each time.

`mk_pkg_from_config()` expects a YAML or JSON file with at least these required
fields:

- `pkg_name`
- `first_name`
- `last_name`

You can also include optional fields like `email`, `git`, `license`,
`check_pkg_name`, `readme_md`, and `pkgdown`.

### Example YAML config

Save a file such as `standard-package.yml`:

```yaml
pkg_name: mypackage
first_name: Jane
last_name: Doe
email: jane.doe@example.com
readme_md: true
git: false
check_pkg_name: true
license: MIT
pkgdown: false
```

This gives you a reusable template. For most new packages, you only need to
change `pkg_name`.

### Create the package from the YAML file

```{r eval = FALSE}
mk_pkg_from_config("standard-package.yml", file_type = "yaml")
```

This helps you keep package setup consistent across projects.

If you want to generate the YAML file from R instead of writing it by hand, use
`write_config()`.

```{r eval = FALSE}
standard_config <- list(
  pkg_name = "mypackage",
  first_name = "Jane",
  last_name = "Doe",
  email = "jane.doe@example.com",
  readme_md = TRUE,
  git = FALSE,
  check_pkg_name = TRUE,
  license = "MIT",
  pkgdown = FALSE
)

write_config("standard-package.yml", standard_config)
```

## Know the built-in validation rules

`pkgmkr` checks common inputs before it creates the package. This helps catch
problems early.

- Package names must start with a letter.
- Package names can contain only letters, numbers, and dots.
- Package names cannot end with a dot.
- Author names must include at least a first and last name.
- Email addresses must be valid if provided.
- Licenses must currently be `MIT` or `GPL-3`.
- The target directory must not already exist.

## Next steps

After the package is created, a common next step is:

1. Add functions in `R/`.
2. Document them with roxygen comments.
3. Run `devtools::document()`.
4. Add tests under `tests/`.
5. Build or refine your pkgdown site if you enabled it.

## Learn more

- Browse the [reference documentation](/reference/) for more details.
- See [examples for `mk_pkg()`](/reference/mk_pkg.html#examples).
- See [configuration helpers](/reference/write_config.html) for the config-file
  workflow.
