---
title: "pkgmkr"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pkgmkr)
```

# Welcome to pkgmkr

`pkgmkr` simplifies the process of creating R packages by providing a single, straightforward function that handles all the setup work for you. Whether you're creating your first package or your hundredth, `pkgmkr` streamlines the process so you can focus on writing code instead of configuration.

## Why pkgmkr?

Creating an R package from scratch involves many steps:
- Setting up the package structure
- Configuring the DESCRIPTION file
- Choosing and setting up a license
- Creating a README
- Initializing Git (if desired)
- Setting up documentation sites

`pkgmkr` automates all of this with one function call, saving you time and ensuring consistency across your packages.

## Quick Start

The simplest way to create a package:

```{r eval = FALSE}
library(pkgmkr)

mk_pkg(
  path = "mypackage",
  author = "Your Name",
  email = "your.email@example.com"
)
```

That's it! `pkgmkr` will create a fully functional R package with:
- Proper package structure
- DESCRIPTION file with your information
- MIT license (or GPL-3 if you prefer)
- README.md file
- Git repository initialized
- Ready for development

## Key Features

### One Function = Complete Setup

`mk_pkg()` handles everything you need to get started:
- Package structure creation
- Metadata configuration
- License setup
- Git initialization (optional)
- README generation
- pkgdown site setup (optional)

### Configuration Files

Prefer configuration files? Use YAML or JSON:

```{r eval = FALSE}
# Create a config file
config <- list(
  pkg_name = "mypackage",
  first_name = "Jane",
  last_name = "Doe",
  email = "jane@example.com",
  license = "MIT"
)

write_config("my_config.yaml", config)
mk_pkg_from_config("my_config.yaml")
```

### Smart Validation

`pkgmkr` includes comprehensive input validation:
- Validates package name format (must follow R conventions)
- Checks email format
- Ensures author names are properly formatted
- Prevents overwriting existing directories
- Validates license choices

### Flexible Options

Customize your package creation:
- Choose between MIT or GPL-3 licenses
- Enable or disable Git initialization
- Control README creation
- Optionally check package name availability on CRAN
- Set up pkgdown documentation site automatically

## What Gets Created?

When you run `mk_pkg()`, you'll get a complete package structure:

```
mypackage/
├── DESCRIPTION          # Package metadata
├── LICENSE             # License file
├── LICENSE.md          # License in markdown
├── NAMESPACE           # Package namespace
├── README.md           # Package README
├── R/                  # Your R code goes here
├── man/                # Documentation (auto-generated)
└── .git/               # Git repository (if enabled)
```

## Next Steps

Once your package is created:

1. **Add your code** to the `R/` directory
2. **Document your functions** using roxygen2 comments
3. **Run `devtools::document()`** to generate help files
4. **Write tests** in the `tests/` directory
5. **Build your site** with `pkgdown::build_site()` (if enabled)

## Learn More

- Browse the [reference documentation](/reference/) for detailed function documentation
- See [examples](/reference/mk_pkg.html#examples) for more usage patterns
- Check out the [configuration guide](/reference/write_config.html) for advanced setups
